---
title: "markdown practice"
author: "Estiaque"
date: "2026-09-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Historical CO2 Analysis

## Introduction

Atmospheric **carbon dioxide (CO~2~)** concentration is an important measure of environmental change. In this exercise, I use the `historic_co2` dataset from the *dslabs* package.

The objective is to examine historical CO~2~ concentrations and convert the original measurement from **parts per million (ppm)** to **parts per billion (ppb)**.

***

## Load the Data

First, I load the required package and dataset.

```{r load-data}
pacman::p_load(dslabs, ggplot2)

data(historic_co2)
```

## Examine the Data

I first examine the names and first few observations of the dataset.

```{r examine-data}
names(historic_co2)

head(historic_co2)

summary(historic_co2)
```

## Data Preparation

The original variable `co2` measures CO~2~ concentration in ppm. I rename this variable `co2_ppm`.

```{r rename-variable}
names(historic_co2)[names(historic_co2) == "co2"] <- "co2_ppm"
```

Because **1 ppm = 1,000 ppb**, I create a new variable called `co2_ppb`.

```{r create-variable}
historic_co2$co2_ppb <- historic_co2$co2_ppm * 1000

head(historic_co2)
```

The mathematical relationship is

$$
CO_{2,ppb}=CO_{2,ppm}\times 1000.
$$

## Variable Description

| Variable | Description | Unit |
|:---------|:-----------:|-----:|
| `year` | Observation year | Year |
| `co2_ppm` | CO~2~ concentration | ppm |
| `co2_ppb` | CO~2~ concentration | ppb |

## Summary Statistics

I calculate several descriptive statistics for atmospheric CO~2~ concentration.

```{r summary-statistics}
mean(historic_co2$co2_ppm)

median(historic_co2$co2_ppm)

sd(historic_co2$co2_ppm)

min(historic_co2$co2_ppm)

max(historic_co2$co2_ppm)
```

## CO~2~ Concentration Over Time

The following figure shows the historical evolution of atmospheric CO~2~ concentration.

```{r co2-graph}
ggplot(historic_co2, aes(x = year, y = co2_ppm)) +
  geom_line(linewidth = 1) +
  labs(
    title = "Historical Atmospheric CO2 Concentration",
    x = "Year",
    y = "CO2 Concentration (ppm)",
    caption = "Source: dslabs"
  ) +
  theme_minimal()
```

## Conclusion

The analysis shows how **R Markdown** can combine *written analysis*, `R commands`, mathematical notation, tables, statistical results, and figures in one document.

~~R code and written analysis must be kept in separate files.~~

Instead, they can be combined in a single reproducible `.Rmd` document.

***